### 2023-02-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@fa8c450e206c47e724b12556ba4c4ee3995ddd4c 🚀

There are a lot of similarities between taxonomic and spatiotemporal
scaling of probabilistic interactions. Fundamentally, these types of
scaling are just different ways to aggregate individuals into broader
nodes, either spatially, temporally, or taxonomically. However, there
are also important differences between them. First, in metawebs, if we
know that two species have the capacity to interact, we can infer that
their respective genus should also be able to interact (i.e., there
should be at least two individuals within these genus that can
interact). On the contrary, knowing that two genus can interact does not
mean that all pairwise combinations of species within these genus can
also interact among themselves. This observation also applies to local
networks. When it comes to probabilistic networks, interaction
probabilities at broader taxonomic scales can be directly obtained from
probabilities at finer scales when aggregating nodes. For example, if we
have in a network \(n_A\) species from genus \(A\) and \(n_B\) species
from genus \(B\), we can calculate the probability that the two genus
interact as
\(P_N(A \rightarrow B) = 1 - \prod_{i = 1}^{n_1}\prod_{j = 1}^{n_2}(1 - P_N(A_i \rightarrow B_j))\),
where \(A_i\) and \(B_j\) are the species of the corresponding genus.
However, more sophisticated models need to be used when building
probabilistic networks at smaller taxonomic resolutions (e.g., when
building a species-level network from a genus-level network). One could,
for example, estimate the probabilities of all pairwise species
interactions by using a Beta distribution parametrised by the
broader-scale network.

>>>>

There are a lot of similarities between taxonomic and spatiotemporal
scaling of probabilistic interactions. Fundamentally, these types of
scaling are just different ways to aggregate individuals into broader
nodes, either spatially, temporally, or taxonomically. However, there
are also important differences between them. First, in metawebs, if we
know that two species have the capacity to interact, we can infer that
their respective genus should also be able to interact (i.e., there
should be at least two individuals within these genus that can
interact). On the contrary, knowing that two genus can interact does not
mean that all pairwise combinations of species within these genus can
also interact among themselves. This observation also applies to local
networks. When it comes to probabilistic networks, interaction
probabilities at broader taxonomic scales can be directly obtained from
probabilities at finer scales when aggregating nodes. For example, if we
have in a network \(n_A\) species from genus \(A\) and \(n_B\) species
from genus \(B\), we can calculate the probability that the two genus
interact as
\(P_N(A \rightarrow B) = 1 - \prod_{i = 1}^{n_A}\prod_{j = 1}^{n_B}(1 - P_N(A_i \rightarrow B_j))\),
where \(A_i\) and \(B_j\) are the species of the corresponding genus.
However, more sophisticated models need to be used when building
probabilistic networks at smaller taxonomic resolutions (e.g., when
building a species-level network from a genus-level network). One could,
for example, estimate the probabilities of all pairwise species
interactions by using a Beta distribution parametrised by the
broader-scale network.
